---
title: "Heart Disease UCI"
author: "Dhiraj Bankar"
date: '2020-Nov-14'
output:
  html_document: default
  word_document: default
  pdf_document: default
bibliography: bibliography.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = '//Users/dhirajbankar/R_Laguage/dsc520/assignments/finalProject/')
```

``` {r include=FALSE , echo=FALSE}
library(readxl)
library(boot)
library(car)
library(mlogit)
library(foreign)
library(caTools)
library(ggplot2)
library(class)
library(Hmisc)
library(Rcmdr)
theme_set(theme_minimal())
```


**Heart Disease UCI Dataset analysis**

``` {r include=FALSE , echo=FALSE}
heartData <- read.csv("heart.csv" )
heartDF <- data.frame(heartData)

```

#### A. Short summary of data
```{r}
str(heartData)
summary(heartData)
```

We have 303 observation with 14 variables:

* age: age
* sex: sex
* cp: chest pain type (4 values)
* trestbps: resting blood pressure
* chol: serum cholesterol in mg/dl
* fbs: fasting blood sugar > 120 mg/dl
* restecg: resting electrocardiograph results (values 0,1,2)
* thalach: maximum heart rate achieved
* exang: exercise induced angina
* oldpeak: ST depression induced by exercise relative to rest
* slope: the slope of the peak exercise ST segment
* ca: number of major vessels (0-3) colored by flourosopy
* thal: 3 = normal; 6 = fixed defect; 7 = reversable defect
* target: Target Groups 

#### B. Find the correleation between all veriables 

``` {r include=TRUE , echo=FALSE}
cor(heartDF)
heartAsMatrix <- as.matrix(heartData[,c("age" ,"trestbps" )])
rcorr(heartAsMatrix)
cor.test(heartDF$age , heartDF$trestbps)
```

From correlation result below are my observations on heart disease data
* The trestbps has 27% affecting to the heart.
* cp  has 27% affecting to the heart.
* Cholesterol has 21% affecting to the heart.
* oldpeakhas 21% affecting to the heart.
* fbs has 12% affecting to the heart.
* exang has 9% affecting to the heart.
* thal has 6% affecting to the heart.
* All other variables has a -ve correlation values(Means they are not impacting or not related to heart disease). We are going to ignore it for further analysis
* For Further analysis lets use only age, trestbps , chol, fbs, exang, oldpeak, ca, thal for main analysis

``` {r include=TRUE , echo=FALSE}
heart_mode_1 <- lm(trestbps ~ chol, data=heartDF)
summary(heart_mode_1)
par(mfrow=c(2,2))
plot(heart_mode_1)
```

``` {r include=TRUE , echo=FALSE}
heart_mode_2 <- lm(trestbps ~ chol+ oldpeak + exang + fbs + ca, data=heartDF)
summary(heart_mode_2)
par(mfrow=c(2,2))
plot(heart_mode_2)
```

``` {r include=FALSE , echo=FALSE}
usingDF1 <- data.frame(heartDF['trestbps'],heartDF['chol'],heartDF['oldpeak'],heartDF['exang'],heartDF['fbs'],heartDF['ca'])
usingDF1$residuals <- resid(heart_mode_2)
usingDF1$standerdized.residuals <- rstandard(heart_mode_2)
usingDF1$studentized.residuals <- rstudent(heart_mode_2)
usingDF1$cooks.distance <- cooks.distance(heart_mode_2)
usingDF1$dfbeta <- dfbeta(heart_mode_2)
usingDF1$dffits <- dffits(heart_mode_2)
usingDF1$leverage <- hatvalues(heart_mode_2)
usingDF1$covariance.ratio <- covratio(heart_mode_2)
usingDF1$large.residual <- usingDF1$studentized.residuals > 2 | usingDF1$studentized.residuals < -2
usingDF1$fitted <- heart_mode_2$fitted.values

```

#### C. What is most common age when heart disease gets started?
``` {r include=TRUE , echo=FALSE}
ggplot(data = heart_mode_1, aes(x = chol )) + geom_histogram(binwidth=10) + labs(x="Age")
```

* From the above histogram I would say heart disease are getting started from thirties of the age

#### D. How much is the effect of cholesterol on heart?
From the correlation found that the Cholesterol has 21% affecting to the heart. Initially before starting the analysis I thought that the Cholesterol is the most important factor which cause the heart disease. But blood pressure and stress has 27% of effect on heart disease. 


``` {r include=TRUE , echo=FALSE}
ggplot(data = heartDF, aes(y = chol, x = trestbps))+
  geom_line()+ geom_smooth(method = "lm" , colour = "Blue" )

scatterplot(chol ~ trestbps , data=heartDF,
            xlab="Blood Pressur", ylab="Cholesterol  Level",
            main="Enhanced Scatter Plot")

```

* The major factor of heart disease is blood pressure and Cholesterol. If you observe above plot, the increase in Cholesterol increases the blood pressure level. 

#### E. Who is most affected by heart disease (Male or Female)? 

``` {r include=TRUE , echo=FALSE}
set.seed(42)
p1 <- hist(heartDF$sex) 
scatterplot(sex ~ trestbps , data=heartDF,
            xlab="Blood Pressur", ylab="Sex",
            main="Enhanced Scatter Plot")

```

* Mail has more heart disease that women. 

#### F. Is fasting blood sugar has a relation with heart disease? 
``` {r include=TRUE , echo=FALSE}
ggplot(data = heartDF, aes(y = trestbps, x = fbs)) + geom_point()+ geom_line(color='red',data = heartDF, aes(y=trestbps, x=fbs))
```

* With fasting the blood pressure increases so the patient with heart disease should not do long fasting. 

#### G. How much help or improvement we can achieve with the help of exercise and diet?
* I couldn't find the exact answer of this question. 

#### H. Will create a multi linear model find the appropriate results. 

``` {r include=TRUE , echo=FALSE}

histogram <- ggplot(usingDF1 , aes (studentized.residuals)) + geom_histogram(aes(y = ..density..), colour = "black" , fill = "white" ) + labs( x= "Studentized Residual", y = "Density") 
histogram + stat_function(fun = dnorm, args = list(mean = mean(usingDF1$studentized.residuals, na.rm = TRUE), sd = sd(usingDF1$studentized.residuals , na.rm = TRUE)), colour = "red", size = 1 ) 

```

``` {r include=TRUE , echo=FALSE}
qplot( sample = usingDF1$studentized.residuals , stat = "qq") + labs( x= "Theoretical Values", y = "Observed Values")
```

``` {r include=TRUE , echo=FALSE}

ggplot(usingDF1 , aes (fitted , studentized.residuals)) + geom_point() + geom_smooth(method = "lm" , colour = "Blue" )  + labs( x= "Fitted Values", y = "Studentized Residual") 

```
``` {r include=TRUE , echo=FALSE}
hist(usingDF1$studentized.residuals)
```


``` {r include=TRUE , echo=FALSE}
hist(rstudent(heart_mode_2))
```


* Overall regression model is unbiased. We could summarize saying that the model appears, in most senses, to be both accurate for the sample and generalizable to the population.

#### I. How much is confidence level of the model?

``` {r include=TRUE , echo=FALSE}
confint(heart_mode_2)
```

* What I observed from the outcome of confident interval. A good model will have a small confidence interval, indicating that the value of b is of the b-values tells us about the direction of the relationship between the predictors and outcome. If you observe outcome of my model is very small, means this model is good. 

## Add Citations
*	Discovering Statistics Using R[@field2012discovering]

# References
